Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2021-12-20 src/ms.tex
jvs comments

For example, efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}. This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with Ro~$>$~2 experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}.

@@

We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}. The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct, though it is interesting that both the long- and short-period pileups seem to be well-described by curves of constant Rossby number. As mentioned above, one theory for the earlier stage of stalled braking is core–envelope coupling. Crucially, in the stalled braking phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core–envelope coupling. In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia. The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.

@@

The data presented here provide evidence for the gap being caused by the same phenomena as the short-period pile-up, namely, core-envelope coupling. The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling. Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling. In this picture, spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability. After some time, spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there. \textbf{An increase in photometric variability accompanies this period of...}

>>>>

For example, efforts to calibrate gyrochronology relations using Kepler asteroseismic targets revealed tension with relations calibrated to open clusters and found that rotation periods could not be described by a single power-law relation with age \citep{Angus2015}. This tension, at least in part, is due to the fact that standard gyrochronology models are unable to account for the anomalously rapid rotation rates of stars older than the Sun, leading to the suggestion that stars with Rossby numbers of Ro~$>$~2 experience a phase of weakened magnetic braking \citep[WMB,][]{vanSaders2016}.

@@

We emphasize that the temporary epoch of stalled magnetic braking proposed by \citet{Curtis2019a, Curtis2020} is not to be confused with the termination of magnetic braking that characterizes the WMB model of \citet{vanSaders2016, vanSaders2019}. The physical mechanisms thought to be responsible for each of these proposed stages of rotational evolution are distinct, though it is interesting that both the long- and short-period pileups seem to be well-described by curves of constant Rossby number. As mentioned above, one theory for the earlier stage of stalled spin-down is core–envelope coupling. Crucially, in the stalled spin-down phase, wind-driven angular momentum losses are not ceased but rather offset by the spin-up torque from core–envelope coupling. In contrast, in the WMB scenario, wind-driven angular momentum losses are proposed to cease ($dJ/dt$=0), with subsequent rotational evolution dictated by the changes in the moment of inertia. The physical mechanism responsible for WMB is not known, but one hypothesis is that the efficiency of magnetic braking declines with magnetic field complexity \citep[e.g.][]{Reville2015, vanSaders2016, Garraffo2016, Metcalfe2016, Metcalfe2019}.

@@

The data presented here provide evidence for the gap being caused by the same phenomena as the short-period pile-up, namely, core-envelope coupling. The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling. Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling. In this picture, spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability. After some time, spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there. \textbf{An increase in photometric variability accompanies this period of post-stalling spin-down, }
